Towards Atomistic Resolution Structure of Phosphatidylcholine Headgroup and Glycerol Backbone at Different Ambient Conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt

Reviewer 2.

We thank for the very positive feedback. The errors in English are
now fixed as suggested.

Reviewer 1.

1.
Author reply:
In the publications mentioned by the reviewer it is shown that the
the membrane properties are affected by the interfacial tension rather than the ensemble. 
This is also explicitly said in the abstract of
J. Chem. Phys. 111, 1281-1287:
"Equivalence between the constant area and constant surface tension ensembles is
investigated by comparing the present simulations with earlier work from our laboratories and we
find simulation results to depend much more strongly on the specified surface area or surface tension
than on the ensemble employed."

In addition, it should be noted that in the mentioned publications the total lengths of the simulations are around 1ns,
while our simulations are 1-2 orders of magnitude longer. 
Thus, even if the dependence on the ensemble would be observed in these short simulations
it would likely be due to insufficient sampling.

As we have already pointed out, we have restricted our work to the available models and
the improvement of these models has been left for further studies. Since all the developers
of the tested models have used NPT ensemble in zero tension, there seems to be consensus on 
this.

It is also much more likely that the force field parameters affect the glycerol backbone and
choline structures much more than the used ensemble or interfacial tension. As we have 
left the more detailed discussion of the force field parameters for further studies, 
the same should be done for the (most likely less relevant) discussion of the ensemble and surface tension.

2.
Author reply:
All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt

To make the definition of the sweet spots clear to the reader, we now explicitly
point this out when introducing them in the text (p15):

"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
subjective sweet spots (light blue areas **spanning 0.04 units around the average
of the extremal experimental values**), within which
we expect the calculated values of the order parameters of a well-performing
force field to fall."

The added words are those between the stars (**).

3.
Author reply:
The title of our manuscript is:
"Towards atomistic resolution structure of
phosphatidylcholine headgroup and glycerol
backbone at different ambient conditions", and
in the introduction we write:
"Phospholipid molecules are composed of hydrophobic acyl chains connected
by a glycerol backbone to a hydrophilic headgroup ..."
This is widely used naming convention for phospholipid molecules.
According to the title and the naming convention used in the manuscript, 
the beginnings of the acyl chains are not the part of headgroup and are 
beyond the scope defined in the title. Their presence at the interface
or their chemical environment do not change this.

We are not willing to extend the scope of the manuscript since it
would significantly increase the already massive information content and 
length. Instead, the acyl chain order parameters will be discussed in 
the context of the NMRLipids III project, see 
http://nmrlipids.blogspot.fi/p/about.html.

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt

To make the definition of the sweet spots clear to the reader, we now explicitly
point this out when introducing them in the text (p15):

"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
subjective sweet spots (light blue areas **spanning 0.04 units around the average
of the extremal experimental values**), within which
we expect the calculated values of the order parameters of a well-performing
force field to fall."

The added words are those between the stars (**).

3.
Author reply:
The title of our manuscript is:
"Towards atomistic resolution structure of
phosphatidylcholine headgroup and glycerol
backbone at different ambient conditions", and
in the introduction we write:
"Phospholipid molecules are composed of hydrophobic acyl chains connected
by a glycerol backbone to a hydrophilic headgroup ..."
This is widely used naming convention for phospholipid molecules.
According to the title and the naming convention used in the manuscript, 
the beginnings of the acyl chains are not part of the headgroup and are 
beyond the scope defined in the title. Their presence at the interface
or their chemical environment do not change this.

### 2015-10-26 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Small updates to the cholesterol and dehydration figure
Add keywords
Update references and fix a typo from https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/68

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

@@

\subsection{Order parameters from experiments}\label{expORDp}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^1$H-$^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions, and also the connections between order parameter and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{h}{e^2qQ} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation 
$|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using the equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7~kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_\mathrm{CH}$ 
through a scaling factor that depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion, see Ref.~\citenum{accuracyPOST}.

@@

We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
effects of cholesterol
on the glycerol backbone order parameters of POPC
measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
This further supports the ideas that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  %Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 for g$_1$ are beyond the y-axis scale.
    In order to elucidate the relative change as a function of cholesterol content, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (S$_\mathrm{CH}$+shift) in the y-label.
}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{lipid, bilayer, molecular dynamics simulation, nuclear magnetic resonance, order parameter}

@@

\subsection{Order parameters from experiments}\label{expORDp}
The order parameter of a hydrocarbon C--H vector is defined as 
\begin{equation}\label{orderP}
S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
\end{equation} 
where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^1$H-$^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions, and also the connections between order parameter and quadrupolar or dipolar splitting
are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{h}{e^2qQ} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation 
$|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using the equation
$|S_{{\rm CH}}|=2\pi\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2--22.7~kHz are used for $(2\pi\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c})^{-1}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_\mathrm{CH}$ 
through a scaling factor that depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For a more detailed discussion, see Ref.~\citenum{accuracyPOST}.

@@

We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
effects of cholesterol
on the glycerol backbone order parameters of POPC
measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
This further supports the ideas that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization.pdf}
%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  %Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 for g$_1$ are beyond the y-axis scale.
    In order to elucidate the relative change as a function of cholesterol content, the simulation results were vertically shifted; 
    the shift magnitudes for each of the force fields are listed (S$_\mathrm{CH}$+shift) in the y-label.
}
\end{figure}
